---
title: "RBioFormats: an R interface to the Bio-Formats library"
author: Andrzej Oleś
package: "`r paste(pkg_ver('RBioFormats'), sprintf('(Bio-Formats library version: %s)', RBioFormats::BioFormats.version()))`"
output:
  BiocStyle::html_document:
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{RBioFormats: an R interface to the Bio-Formats library}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
library("knitr")
```

```{r RBioFormats, message=FALSE, echo=FALSE}
library("RBioFormats")
```


# Reading images

```{r library, message=FALSE}
library("RBioFormats")
```

Read a greyscale image

```{r readgrey}
f = system.file("images", "sample.png", package = "EBImage")

img = read.image(f)
img
```

or an RGB image

```{r readrgb}
f = system.file("images", "sample-color.png", package = "EBImage")

img = read.image(f)
img
```

Object summary without the image data preview can be printed using the `short = TRUE` argument to `print`.

```{r printshort}
print(img, short = TRUE)
```

For displaying only the dimension order use

```{r dimorder}
dimorder(img)
```


# The *AnnotatedImage* class

`r Rpackage('RBioFormats')` stores image data in an *AnnotatedImage* class which extends the *Image* class from `r Biocpkg('EBImage')`.

```{r classdef}
getClassDef("AnnotatedImage")
```

Compared to the original *Image* class the *AnnotatedImage* class features an additional `metadata` slot containing image metadata.

```{r metadata}
meta = metadata(img)
meta
```

To alter the length of the printed output use the `list.len` attribute to `print`:

```{r printmeta}
print(meta, list.len=99L)
```


# Image metadata

Image metadata is represented by an *ImageMetadata* class structured as a named list of coreMetadata, globalMetadata and seriesMetadata.

```{r metaNames}
names(meta)
cMeta = meta$coreMetadata
names( cMeta )
```

`coreMetadata` stores information which is guaranteed to exist for any image type, whereas the latter two metadata types are format-specific and can be empty. 

Each of these metadata sublists has an correspodning accessor function, e.g.,

```{r coreMetadata}
identical( coreMetadata(meta), cMeta)
```

and similarly for `globalMetadata` and `seriesMetadata`. These accessors are useful for extracting the corresponding metadata directly from an *AnnotatedImage* object

```{r coreMetadata2}
identical( coreMetadata(img), cMeta )
```


# Reading image metadata

Image metadata can be read without loading the corresponding pixel data.

```{r read.metadata}
f = system.file("images", "nuclei.tif", package = "EBImage")
metadata = read.metadata(f)

metadata
```

This is especially useful for large data sets.


# OME-XML representation

and the OME-XML DOM tree representation of the metadata in the `omexml` slot.

The OME-XML metadata can be accessed using tools from the `r CRANpkg('XML')` package.

```{r parseXML, message=FALSE, results='hide'}
require("XML")

omexml = xmlTreeParse(read.omexml(f))
xmlRoot(omexml)
```

For details on working with XML data in R see the `r Rpackage('XML')` package [documentation](http://www.omegahat.org/RSXML/ "XML package documentation").


# Session info {.unnumbered}

Here is the output of `sessionInfo()` on the system on which this
document was compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# Appendix A: Working with test images {.unnumbered}

For development purposes it is useful to have images of a specific size or pixel type for testing. Mock files containing gradient images can be generated using

```{r mockFile, out.width='256px', out.height='256px'}
f = mockFile(sizeX = 256, sizeY = 256)
img = read.image(f)

require("EBImage")
display(img, method = "raster")
```

Note that the native image data range is different depending on pixel type:

```{r defaultRange, echo=FALSE, results='asis', R.options=list(digits = 15)}
FormatTools = J("loci.formats.FormatTools")
types = c("int8", "uint8", "int16", "uint16", "int32", "uint32", "float", "double")

ranges = sapply(types, function(t) {
  minmax = FormatTools$defaultMinMax(FormatTools$pixelTypeFromString(t))
  setNames(minmax, c("min", "max"))
  })
knitr::kable(ranges)
```

Image data returned by `r Rpackage('RBioFormats')` is by default scaled to the [0:1] range. This behavior can be controlled using the `normalize` argument to `read.image`.

```{r range}
sapply(types, function(t) {
  img = read.image(mockFile(sizeX = 65536, sizeY = 11, pixelType = t), normalize = FALSE)
  if (typeof(img)=="raw") img = readBin(img, what = "int", n = length(img), size = 1L)
  setNames(range(img), c("min", "max"))  
})
```

# Appendix B: Compared to *EBImage* {.unnumbered}

Loading images using `r Githubpkg('aoles/RBioFormats')` should give the same results as using the `r Biocpkg('EBImage')` package.

```{r comparewithref}
require("EBImage")
f = system.file("images", "sample-color.png", package = "EBImage")
identical(readImage(f), as.Image(read.image(f)))
```

Performance relative to *EBImage*

```{r benchmark, message=FALSE, eval=FALSE}
require("microbenchmark")
microbenchmark(read.image(f), readImage(f), times = 10)
```


